---
title: "1_dataclean_bg"
author: "Karin Emanuelson"
date: "12/2/2019"
output: html_document
bibliography: library.bib
editor_options: 
  chunk_output_type: console
---
# Calculating Uptake: Clean Background Data (step 1 of 3)

## Study Background

A series of four-hour nutrient enrichment injections were conducted at Como Creek, CO during the summer of 2018 as part of a collaborative NSF funded research project that is evaluating the hydrologic and stoichiometric conditions that maximize N retention. Each series of enrichment injections consisted of the following treatments:

1. nitrogen alone, 
2. nitrogen and carbon, 
3. nitrogen and phosphorus, and 
4. nitrogen, carbon and phosphorus 
* all co-injected with a conservative tracer (chloride). 

Grab samples were taken throughout the experiments to determine background concentrations, plateau concentrations and concentrations down the falling limb of the breakthroughcurve. This analysis will determine the uptake length, average distance traveled by a dissolved nutrient before biotic uptake, for nitrogen for each of the four treatments [@Tank2017].

```{r setup, warning = FALSE, echo = TRUE, message=FALSE}
## Setup
knitr::opts_knit$set(root.dir='./..')
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message=FALSE)
library(tidyverse)
library(lubridate)
library(xts)
library(dygraphs)
library(ggrepel)
library(knitr)
library(plotly)
library(ggthemes)
library(RColorBrewer)
library(htmltools)
```

## Data Cleaning: Background Samples

Background longitudinal samples were taken every 50 meters through the stream reach each day before the nutrient enrichment injections began. These samples represent the ambient nutrient concentration in the stream. All data will need to be background corrected in order to only account the for the added nutrients in the analysis. As the first step, raw chemistry data is read in, filtered to only include the background samples, and flagged for minimum and maximum detection limits.

```{r}
## Load Raw Chem Data for Background Samples only
chem_grab_bg<- read_csv(('data/in/2018_ComoCreek_WaterChemistry_Mastersheet_Recreated_110519_SORTED.csv'), skip=7)%>%
      mutate(Datetime= mdy_hm(Datetime,tz='MST7MDT'))%>%
      mutate(run_date= mdy(run_date))%>%
      select(-ID)%>%
      mutate(id = rownames(.))%>%
  select(
    sample = "Sample #" ,
    id,
    site = Location,
    datetime = Datetime,
    compartment = Type1,
    sample_type = Type2,
    rundate = run_date,
    df = DF,
    Na:SO4_DIL
  )%>%
    filter(compartment == 'MC') %>%
  filter(sample_type =='BG LONG')

# lookup table for chem Max/Min
chem_qc= data.frame(var = c('Cl','NO3','PO4'), MaxDL=c(1.7, 5, 7.5), MDL=c(.01, .01, .01))

```


```{r}
### Create long dataset
chem_bg<-chem_grab_bg %>%
  select(sample:SO4)%>%
  gather(var, value, Na:SO4) %>%
  filter(var %in% c('NO3','Cl','PO4'))%>%
  mutate(value = as.numeric(value))%>%
  mutate(site = as.numeric(site))%>%
  mutate(df = if_else(is.na(df),1,df))%>%
  mutate(conc_type = 'final') %>%
  mutate(value = if_else(is.na(value), .005, value))

## Flag for minimum and maximum detection limits
chem_flag<- 
  chem_bg%>%
  mutate(value = value/df)%>%
  mutate(conc_type='raw')%>%
  left_join(chem_qc)%>%
  mutate(MaxDL_flag = if_else(value>MaxDL, 'Yes','No'))%>%
  mutate(MDL_flag= if_else(value<MDL, 'Yes','No'))%>%
  select(id, var, MaxDL_flag, MDL_flag)

## Add flags to long dataset  
chem_long_bg<-
  chem_bg%>%
  left_join(chem_flag)%>%
  mutate(id = rownames(.))
```


```{r}
### Define function to graph data and set color scheme
plotChem <- function (df, choose_param=c( 'Cl', 'NO3', 'PO4')) {
  date_title<-date(df$datetime[[1]]) #Define title for each graph
  ggplotly(
    ggplot(
      dplyr::filter(df, var %in% choose_param),
      aes(
        site,
        value,
        fill = c(sample_type),
        text = paste('date: ',datetime, #Define pop-up info
          '<br> value:',value,
          '<br> sample_type:',sample_type,
          '<br> rundate:',rundate,
          '<br> sample:',sample,
          '<br> df:',df,
          '<br> MaxDL_flag:',MaxDL_flag,
          '<br> MDL_flag:',MDL_flag,
          '<br> id:',id
        ),
        group = 1
      )
    ) +
      geom_point(shape = 21, size = 3) +
      geom_line() +
      theme_few() +
      scale_fill_manual(values = col_vector) +
      facet_wrap( ~ var, ncol = 1, scale = 'free_y')+
      labs(title = date_title),
    width = 1200,
    height =    1000,
    tooltip = 'text'
  )
}

#create color scheme for chem plots
n <- 40
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

```

### Visualize Raw Data

Graph background longitudinal sample concentrations for the ions of interest (NO3, PO4, and Cl) each day. Visually inspect data for outliers, and create "good id" vector of chosen datapoints. Hover over graphed points to see sample ID, dilution factor, rundate, min and max flags, and concentration value.

```{r}
# Filter for each injection
chem_grab_nest_bg<-chem_long_bg%>%
  mutate(date=date(datetime))%>%
  arrange(date)%>%
  group_by(date)%>%
  nest()

bg<-map(chem_grab_nest_bg$data, plotChem)

bg[[1]]
bg[[2]]
bg[[3]]
bg[[4]]
bg[[5]]
bg[[6]]
bg[[7]]
bg[[8]]

```

#### Create good_id vector and only select those datapoints. Check count of flagged data. If count is not 0, inspect dataset to ensure validity of chosen data. 
```{r}

good_id_bg<-c(1:40,43:81,82:105,108:110,112:121,124:136,138:162,163:202,205:243)

#create dataset of only selected points
chem_clean_long_bg <- chem_long_bg%>%
  filter(id %in% good_id_bg)%>%
  mutate(date=date(datetime))

#double check that Max Detection Limit is vaild
chem_clean_long_flagged<-chem_clean_long_bg%>%
  filter(MaxDL_flag== 'Yes')

count(chem_clean_long_flagged)
```

### Visualize Selected Data

Graph selected background longitudinal sample concentrations for the ions of interest (NO3, PO4, and Cl) each day. Use this as a double check before you continue to use these data points in future analysis
```{r}
# Double check everything
chem_clean_nest_bg<-chem_clean_long_bg%>%
  arrange(date)%>%
  group_by(date)%>%
  nest()


c<-map(chem_clean_nest_bg$data, plotChem)

c[[1]]
c[[2]]
c[[3]]
c[[4]]
c[[5]]
c[[6]]
c[[7]]
c[[8]]
```

## Summarize Data
Take the mean of the sample concentrations for each ion of interest (NO3, PO4, and Cl) each day. This background mean concentration will be used to background correct plateau concentrations.
```{r}
chem_mean_bg<- chem_clean_long_bg%>%
  group_by(date, var)%>%
  summarize(bg_mean=mean(value))

background_means<-chem_mean_bg%>%
    spread(var, bg_mean)
kable(background_means)

write_csv(background_means,"data/in/background_means.csv" )
write_csv(chem_mean_bg,"data/in/chem_mean_bg.csv" )

```




